Estimating the dynamics of SPARQL query results

### 2019-04-11 kgswc19.tex
update

One topic we do wish to highlight however -- as it relates to the behavior of a query over a dynamic RDF graph -- is that of the \emph{monotonicity} of SPARQL queries~\cite{ArenasP11,ArenasU17}. We say that a \texttt{SELECT} query $Q$ is monotone if and only if $G_1 \subseteq G_2$ implies that $Q(G_1) \subseteq Q(G_2)$ for RDF graphs $G_1$ and $G_2$; intuitively, as data are extended, the results of a monotone query can only be extended. Monotonic SPARQL features include, for example, joins, unions, paths and filters; on the other hand, non-monotonic SPARQL features include negation and optional~\cite{ArenasP11}.

@@

The second block indicate the query operators used; we capture the presence or absence of query operators and solution modifiers from SPARQL 1.1. In the third block, we group related features into one dimension: in the case of \textit{Recursive path}, we group queries that contain as expression of the form $e$\texttt{*} or $e$\texttt{+}. On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{!BOUND}). These features -- though course-grained -- are straightforward to extract from a query, and offer valuable insights into how the query may behave in a dynamic query; for example, the \textit{Negation} feature captures information about the (non-)monotonicity of the query, while we suppose that \textit{Recursive paths}, which may traverse an arbitrary number of triples in the graph, might be more sensitive to change. Again, such correlations are not without exception and will require experimental scrutiny.

@@

\paragraph{RDF Data:} We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured almost weekly($\pm$ 1 day) in the \emph{truthy version} that contains triples (without qualifiers) that have the best non-deprecated rank for a given property. The first version has 1,102,242,331 triples and 3,276 unique predicates, while the final version has 1,924,967,162 triples (+74\%) and 3,660 unique predicates (+11\%). Figure~\ref{fig:wikidata}, on the left, shows the growth in triples as the time progress, while on the right, shows the numbers of triples added and removed version-to-version. We see that although many more triples are added than removed, there are some triples removed each version. We further note some peaks in triples added in some versions, which may be due to bulk imports of data. Between the version 11 and 12 we have almost 2 weeks because we were not able to obtain the data for that week, but this cause the third highest peak. The dynamic graph thus considers a total of 32.3 billion triples across 23 versions.

@@

\paragraph{Setting:} To build our final datasets, given that our data features are temporal based, we must thus define a past interval to consider. With 23 versions, we must hold out at least one version to label, leaving a maximum interval of 22 past versions to use for compute our data features. However, the more past versions we consider for a prediction, the fewer examples we can generate; for example, if we consider 22 versions for compute, we will only have one ground truth label for each query. In the end, we opted to consider intervals of 3, 5, 9 and 17 previous versions, for example, ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes in queries results for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query and 4420 examples in the largest dataset and 1326 examples in the smallest dataset. We remark that no feature explicitly indicates which specific version numbers are given and which are to be predicted.

@@

For OSC prediction, we selected four classifiers for testing: Decision Trees, Linear SVM, Naive Bayes and Nearest Neighbors. We split the data to use 80\% for training and 20\% for tests. To avoid overfitting, we use stratified cross-validation in the training data. We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) considering information of $\Delta(\mathcal{G},p)$ and the cardinality of predicates and information on historical results (\textsc{r}) considering the number of past changes in the fixed interval. We may also consider combinations of these features: \textsc{qp}, \textsc{qr}, \textsc{pr} and \textsc{qpr}.

@@

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.774)(5,0.787)(3,0.796)(4,0.815)
	};

@@

Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} library. The latter allows us to see how the periods of estimation and the convergence analysis will behave. We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 weeks and a variance score of 0.01. These results allow to predict future changes in query results with good accuracy using linear regression techniques.

@@

Our goal then was to explore this trade-off experimentally. For this, we define some desirable characteristics for the queries and data used in the experimental setting, such as availability, size, dynamics, diversity, etc. Settling on Wikidata, we collect 23 versions and curate 221 user-generated queries from which we compute results, statistics and ultimately features that form the gold standard for the tasks. Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus a non-trivial task. We make this gold standard available at \ah{...} -- including queries, results and features -- so that the results presented here can be reproduced and so that it can be reused in further research.

>>>>

One topic we do wish to highlight, however -- as it relates to the behavior of a query over a dynamic RDF graph -- is that of the \emph{monotonicity} of SPARQL queries~\cite{ArenasP11,ArenasU17}. We say that a \texttt{SELECT} query $Q$ is monotone if and only if $G_1 \subseteq G_2$ implies that $Q(G_1) \subseteq Q(G_2)$ for RDF graphs $G_1$ and $G_2$; intuitively, as data are extended, the results of a monotone query can only be extended. Monotonic SPARQL features include, for example, joins, unions, paths and filters; on the other hand, non-monotonic SPARQL features include negation and optional~\cite{ArenasP11}.

@@

The second block indicates the query operators used; we capture the presence or absence of query operators and solution modifiers from SPARQL 1.1. In the third block, we group related features into one dimension: in the case of \textit{Recursive path}, we group queries that contain as expression of the form $e$\texttt{*} or $e$\texttt{+}. On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{!BOUND}). These features -- though course-grained -- are straightforward to extract from a query, and offer valuable insights into how the query may behave in a dynamic query; for example, the \textit{Negation} feature captures information about the (non-)monotonicity of the query, while we suppose that \textit{Recursive paths}, which may traverse an arbitrary number of triples in the graph, might be more sensitive to change. Again, such correlations are not without exception and will require experimental scrutiny.

@@

\paragraph{RDF Data:} We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured almost weekly($\pm$ 1 day) in the \emph{truthy version} that contains triples (without qualifiers) that have the best non-deprecated rank for a given property. The first version has 1,102,242,331 triples and 3,276 unique predicates, while the final version has 1,924,967,162 triples (+74\%) and 3,660 unique predicates (+11\%). Figure~\ref{fig:wikidata}, on the left, shows the growth in triples as the time progress, while on the right, shows the numbers of triples added and removed version-to-version. We see that although many more triples are added than removed, there are some triples removed each version. We further note some peaks in triples added in some versions, which may be due to bulk imports of data. Between version 11 and 12 we have almost 2 weeks because we were not able to obtain the data for that week, but this cause the third highest peak. The dynamic graph thus considers a total of 32.3 billion triples across 23 versions.

@@

\paragraph{Setting:} To build our final datasets, given that our data features are temporally based, we must thus define a past interval to consider. With 23 versions, we must hold out at least one version to label, leaving a maximum interval of 22 past versions to use to compute our data features. However, the more past versions we consider for a prediction, the fewer examples we can generate; for example, if we consider 22 versions to compute, we will only have one ground truth label for each query. In the end, we opted to consider intervals of 3, 5, 9 and 17 previous versions, for example, ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes in queries results for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query and 4420 examples in the largest dataset and 1326 examples in the smallest dataset. We remark that no feature explicitly indicates which specific version numbers are given and which are to be predicted.

@@

For OSC prediction, we selected four classifiers for testing: Decision Trees, Linear SVM, Naive Bayes, and Nearest Neighbors. We split the data to use 80\% for training and 20\% for tests. To avoid overfitting, we use stratified cross-validation in the training data. We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) considering information of $\Delta(\mathcal{G},p)$ and the cardinality of predicates and information on historical results (\textsc{r}) considering the number of past changes in the fixed interval. We may also consider combinations of these features: \textsc{qp}, \textsc{qr}, \textsc{pr} and \textsc{qpr}.

@@

@@

Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} library. The latter allow us to see how the periods of estimation and the convergence analysis will behave. We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 weeks and a variance score of 0.01. These results allow us to predict future changes in query results with good accuracy using linear regression techniques.

@@

Our goal then was to explore this trade-off experimentally. For this, we define some desirable characteristics for the queries and data used in the experimental setting, such as availability, size, dynamics, diversity, etc. Settling on Wikidata, we collect 23 versions and curate 221 user-generated queries from which we compute results, statistics and ultimately features that form the gold standard for the tasks. Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus a non-trivial task. We make this gold standard available at \ah{...} -- including queries, results, and features -- so that the results presented here can be reproduced and so that it can be reused in further research.
